Supporting Information: Unsupervised embedding of trajectories captures the latent structure of mobility

### 2021-03-10 paper/current/supp.tex
supp text tweak

Why do neural-embedding approaches, which rely on stochastic gradient descent, outperform Levy's direct matrix factorization~\autocite{levy2014neural}, especially given that \textit{word2vec} \emph{is} implicitly approximating factorization?
We speculate that is stems, in part, from the sensitivity of matrix factorization to small flows between locations.  
Levy's matrix factorization embeds affiliations $i$ and $j$ such that their dot similarity is as close as possible to $\log (T_{ij} / P(i) P(j))$. 
If the flow $T_{ij}$ is considerably small or zero, the dot similarity goes to $-\infty$, pushing $i$ and $j$ very far from other affiliations in the embedding space~\autocite{levy2014neural, Qui2018}. 
This is particularly problematic when the window size is small because most affiliation pairs would have no flow, which is indeed the case in our experiments. 
To circumvent this problem, previous studies~\autocite{levy2014neural, Qui2018} added a constant flow between the affiliation pairs with no flow. 
However, this alters the underlying data, and these small flows can still have a strong impact on the embedding.

@@

%
% S11 TEXT
\paragraph*{S11 Text}
\label{si:text:disambiguation}
{\bf Author name dismabiguation.}

>>>>

Why do neural-embedding approaches, which rely on stochastic gradient descent, outperform Levy's direct matrix factorization~\autocite{levy2014neural}, especially given that \textit{word2vec} \emph{is} implicitly approximating factorization?
We speculate that is stems, in part, from the sensitivity of matrix factorization to small flows between locations.  
Levy's matrix factorization embeds affiliations $i$ and $j$ such that their dot similarity is as close as possible to $\log (T_{ij} / P(i) P(j))$. 
If the flow $T_{ij}$ is considerably small or zero, the dot similarity goes to $-\infty$, pushing $i$ and $j$ very far from other affiliations in the embedding space~\autocite{levy2014neural, Qui2018}. 
This is particularly problematic when the window size is small because most affiliation pairs would have no flow, which is indeed the case in our experiments. 
To circumvent this problem, previous studies~\autocite{levy2014neural, Qui2018} added a constant flow between the affiliation pairs with no flow. 
However, in addition to altering the underlying data, these small flows can still have a strong impact on the embedding.

@@

%
% S11 TEXT
\paragraph*{S11 Text}
\label{si:text:disambiguation}
{\bf Author name disambiguation.}
